---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# did

The \code{did} package contains tools for computing average treatment effect parameters in Difference in Differences models with more than two periods and with variation in treatment timing.  The main parameters are group-time average treatment effects which are the average treatment effect for a particular group at a a particular time.  These can be aggregated into a fewer number of treatment effect parameters, and the package deals with the cases where there is selective treatment timing, dynamic treatment effects, calendar time effects, or combinations of these.  There are also functions for testing the Difference in Differences assumption, and plotting group-time average treatment effects.

## Installation

You can install \texttt{did} from CRAN with:

```{r eval=FALSE}
install.packages("did")
```

or get the latest version from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("bcallaway11/did")
```

## Example

The following is a simplified example of the effect of states increasing their minimum wages on county-level teen employment rates which comes from Callaway and Sant'Anna (2018).

A subset of the data is available in the package and can be loaded by
```{r}
  library(did)
  data(mpdta)
```

The dataset contains 500 observations of county-level teen employment rates from 2003-2007.  Some states are first treated in 2004, some in 2006, and some in 2007 (see the paper for details).  We are interested in estimating the group-time average treatment effect, which is the average treatment effect in period $t$ for the group of states first treated in period $g$ and given by
\begin{align*}
ATT(g,t)=E[Y_t(1) - Y_t(0)|G_g=1]
\end{align*}
under the common trends assumption:
\begin{align*}
  E[\Delta Y_t(0) | X, G_g=1] = E[\Delta Y_t(0) | X, C=1] \ a.s. \quad \forall g\leq t
\end{align*}
where $Y_t(1)$ and $Y_t(0)$ denote treated and untreated potential outcomes, $G_g=1$ denotes counties first treated in period $g$, $C=1$ denotes control counties that are never treated.

To estimate $ATT(g,t)$, one can use the \texttt{mp.spatt} method as follows:

```{r}
out <- mp.spatt(lemp ~ treat, xformla=~lpop, data=mpdta,
                 panel=TRUE, first.treat.name="first.treat",
                 idname="countyreal", tname="year",
                 bstrap=FALSE, se=TRUE, cband=FALSE)
summary(out)
```
and plot the results using the \texttt{ggdid} command:

```{r}
library(gridExtra)
ggdid(out)
```


