---
title: "DID Basics"
author: "Brantly Callaway and Pedro H.C. Sant\'Anna"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DID Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette discusses the basics of using Difference-in-Differences (DID) designs to identify and estimate the average effect of participating in a treatment with a particular focus on tools from the `did` package.

# A Running Example

Throughout the vignette, we use a subset of data that comes from Callaway and Sant'Anna (2019).  This is a dataset that contains county-level teen employment rates from 2001-2007.  The data can be loaded by
```{r}
library(did)
##data(mpdata)
```
And the dataset looks like
```{r}
##head(mpdata)
```

## Some Comments about the Format of the Data for using the 'did` Package



# Identification

First, we provide a brief overview of how identification works as well as parameters of interest in DID designs.

The main identifying assumption

# Estimation

## Two-Groups / Two Periods

## Multiple Groups and Periods

# Common Issues using the `did` package

# References 