---
title: "Getting Started with the did Package"
author: "Brantly Callaway and Pedro H.C. Sant&apos;Anna"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the did Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
fldr <- "~/Dropbox/did/R/"
sapply(paste0(fldr,list.files(fldr)), source)
library(DRDID)
library(BMisc)
library(ggplot2)
library(ggpubr)
```


This vignette discusses the basics of using Difference-in-Differences (DID) designs to identify and estimate the average effect of participating in a treatment with a particular focus on tools from the `did` package.

# Examples with simulated data

Let's start with a really simple example with simulated data.  Here, there are going to be 4 time periods.  There are 4000 units in the treated group that are randomly (uniformly) assigned to first participate in the treatment (a *group*) in each time period.  And there are 4000 never treated units.  The data generating process for untreated and treated potential outcomes is given by

$$
  Y_{it}(0) = \theta_t + \eta_i + X_i'\beta_t + v_{it}
$$

$$
  Y_{it}(1) = \theta_t + \eta_i + X_i'\beta_t + u_{it}
$$

where $\theta_t = \beta_t = t$ for $t=1,\ldots,4$, $\eta_i \sim N(G_i, 1)$ where $G_i$ is which group an individual belongs to, $X_i \sim N( \mu_{D_i}, 1)$ where $\mu_{D_i} = 1$ for never treated units and 0 otherwise, $v_{it} \sim N(0,1)$, and $u_{it} \sim N(0,1)$.  Also, all random variables are drawn independently of each other.  In this setup $ATT(g,t) = 0$ for all groups and time periods.

## Estimating Group-Time Average Treatment Effects

**Building the dataset**

```{r echo=FALSE}
source("~/Dropbox/did/tests/setup_sims.R")
time.periods <- 4
reset.sim()
te <- 0
set.seed(1814)
```

```{r}
# generate dataset with 4 time periods
time.periods <- 4

# generate data set with these parameters
data <- build_sim_dataset()

nrow(data)
head(data)
```

**Estimating Group-Time Average Treatment Effects**

The main function to estimate group-time average treatment effects is the `att_gt` function.  [See documentation here](reference/att_gt.html).  The most basic call to `att_gt` is given in the following

```{r}
# estimate group-time average treatment effects using att_gt method
example.attgt <- att_gt(yname="Y",
                        tname="period",
                        idname="id",
                        first.treat.name="G",
                        xformla=~X,
                        data=data,
                        printdetails=FALSE
                        )

# summarize the results
summary(example.attgt)
```

The summary of `example.attgt` provides estimates of the group-time average treatment effects in the column labeled `att`, and the corresponding standard errors are given in the column `se`.  The corresponding groups and times are given in the columns `group` and `time`.  Under the parallel trends assumption, group-time average treatment effects are identified in periods when $g \leq t$ (i.e., post-treatment periods for each group).  The table also reports pseudo group-time average treatment effects when $g > t$ (i.e., pre-treatment periods for group $g$).  These can be used as a pre-test for the parallel trends assumption.  In addition, the results of a Wald pre-test of the parallel trends assumption is reported in the summary of the results.  [A much more detailed discussion of using the **did** package for pre-testing is available here](pre-testing).

Next, we'll demonstrate how to plot group-time average treatment effects.  To plot these, use the `ggdid` function which builds off the `ggplot2` package.
```{r}
# plot the results
ggdid(example.attgt)
```

The resulting figure is one that contains separate plots for each group.  Notice in the figure above, the first plot is labeled "Group 2", the second "Group 3", etc.  Then, the figure contains estimates of group-time average treamtent effects for each group in each time period along with a confidence interval.  The red dots in the plots are pre-treatment pseudo group-time average treatment effects and are most useful for pre-testing the parallel trends assumption.  The blue dots are post-treatment group-time average treatment effects and should be interpreted as the average effect of participating in the treatment for units in a particular group at a particular point in time.

## Extensions

The above discussion covered only the most basic case for using the `did` package.  There are a number of extensions that are useful in applications.  

### Adjustments for Multiple Hypothesis Testing

By default, the `did` package reports analytical standard errors and 95\% *pointwise* confidence intervals in plots.  However, it is usually a better practice to report uniform confidence bands in plots of group-time average treatment effects with multiple time periods -- these are confidence bands that are robust to multiple hypothesis testing.

To report uniform confidence bands requires the bootstrap; and to get these in practice, set the options `bstrap=TRUE, cband=TRUE`.

```{r eval=FALSE}
  example.attgt.mh <- att_gt(yname="Y",
                        tname="period",
                        idname="id",
                        first.treat.name="G",
                        xformla=~X,
                        data=data,
                        printdetails=FALSE,
                        bstrap=TRUE,
                        cband=TRUE
                        )
  ## plot (not shown, but will result in somewhat wider confidence bands)
  ggdid(example.attgt.mh)
```

### Aggregating group-time average treatment effects

In many applications, there can be a large number of groups and time periods.  In this case, it may be infeasible to interpret plots of group-time average treatment effects.  The `did` package provides a number of ways to aggregate group-time average treatment effects using the `aggte` method.

#### Simple Aggregation

The default approach is to just return a weighted average of all group-time average treatment effects with weights proportional to the group size.

```{r}
agg.simple <- aggte(example.attgt)
summary(agg.simple)
```

#### Dynamic Effects and Event Studies

The most common alternative approach is to aggregate group-time effects into an event study plot.  To do this

```{r}
agg.es <- aggte(example.attgt, type="dynamic")
summary(agg.es)
ggdid(agg.es)
```

In this figure, the x-axis is the length of exposure to the treatment.  Length of exposure equal to 0 provides the average effect of participating in the treatment across groups in the time period when they first participate in the treatment.  Length of exposure equal to -1 corresponds to the time period before groups first participate in the treatment, and length of exposure equal to 1 corresponds to the first time period after initial exposure to the treatment.

Not much interesting is going on here...as there are no dynamics and the effect is equal to 0 in all time periods.

#### Selective Treatment Timing / Group-Specific Effects

Another idea is to look at average effects specific to each group.  To do this
```{r}
agg.gs <- aggte(example.attgt, type="selective")
summary(agg.gs)
ggdid(agg.gs)
```

In this figure, the x-axis is categorized by group.  They y-axis provides estimates of the average effect of participating in the treatment for units in each group averaged across all time periods after that group becomes treated.

#### Calendar Time Effects

Finally, the `did` package allows aggregations across different time periods.  To do this

```{r}
agg.ct <- aggte(example.attgt, type="calendar")
summary(agg.ct)
ggdid(agg.ct)
```

In this figure, the x-axis is the time period and the estimates along the y-axis are the average effect of participating in the treatment in a particular time period for all groups that participated in the treatment in that time period.

### Small Group Sizes

Small group sizes can sometimes cause estimation problems in the `did` package.  To give an example, if there are any groups that have fewer observations than the number of covariates in the model, the code will error.  The `did` package reports a warning if there are any groups that have fewer than 5 observations.

In addition, statistical inference, particularly on group-time average treatment may become more tenuous with small groups.  For example, the effective sample size for estimating the change in outcomes over time for individuals in a particular group is equal to the number of observations in that group and asymptotic results are unlikely to provide good approximations to the sampling distribution of group-time average treatment effects when the number of units in a group is small.

A reasonable alternative approach in this case is to just focus on aggregated treatment effect paramaters (i.e., to run `aggte(..., type="selective")` or `aggte(...,type="dynamic")`).  For each of these cases, the effective sample size is the total number of units that are ever treated.  As long as the total number of ever treated units is "large" (which should be the case for any DID application), then the statistical inference results provided by the `did` package should be reliable.

### Selecting Alternative Control Groups

By default, the `did` package uses the group of units that never participate in the treatment as the control group.  In this case, if there is no group that never participates, then the `did` package will drop the last period and set units that do not become treated until the last period as the control group (this will also throw a warning).  The other option for the control group is to use the "not yet treated".  The "not yet treated" include the never treated as well as those units that, for a particular point in time, have not been treated yet (though they eventually become treated).  This group is at least as large as the never treated group though it changes across time periods.  To use the "not yet treated" as the control, set the option `control.group="notyettreated"`.

```{r eval=FALSE}
example.attgt.altcontrol <- att_gt(yname="Y",
                        tname="period",
                        idname="id",
                        first.treat.name="G",
                        xformla=~X,
                        data=data,
                        printdetails=FALSE,
                        control.group="notyettreated"			
                        )
summary(example.attgt.altcontrol)
```

### Repeated cross sections

The `did` package can also work with repeated cross section rather than panel data.  If the data is repeated cross sections, simply set the option `panel=FALSE`.  In this case, `idname` is also ignored.  From a usage standpoint, everything else is identical to the case with panel data.

### Alternative Estimation Methods

The `did` package implements all the $2 \times 2$ DID estimators that are in the `DRDID` package.  By default, the `did` package uses "doubly robust" estimators that are based on first step linear regressions for the outcome variable and logit for the generalized propensity score.  The other options are "ipw" for inverse probability weighting and "reg" for regression.

```{r eval=FALSE}
example.attgt.reg <- att_gt(yname="Y",
                        tname="period",
                        idname="id",
                        first.treat.name="G",
                        xformla=~X,
                        data=data,
                        printdetails=FALSE,
                        estMethod="reg"
                        )
summary(example.attgt.reg)
```

The argument `estMethod` is also available to pass in a custom function for estimating DID with 2 periods and 2 groups.  [See its documentation for more details](references/att_gt).  **Fair Warning:** this is very advanced use of the `did` package and should be done with caution.

# An example with real data

Next, we use a subset of data that comes from Callaway and Sant'Anna (2019).  This is a dataset that contains county-level teen employment rates from 2003-2007.  The data can be loaded by
```{r}
library(did)
data(mpdta)
```
`mpdta` is a balanced panel with `r nrow(mpdta)` observations.  And the dataset looks like
```{r}
head(mpdta)
```

**Data Requirements**

In particular applications, the dataset should look like this with the key parts being:

* The dataset should be in *long* format -- each row corresponds to a particular unit at a particular point in time.  Sometimes panel data is in *wide* format -- each row contains all the information about a particular unit in all time periods.  To convert from wide to long in `R`, one can use the `tidyr::gather` function.  [Here is an example](http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/) 

* There needs to be an id variable.  In `mpdta`, it is the variable `countyreal`.  This should not vary over time for particular units.  The name of this variable is passed to methods in the `did` package by setting, for example, `idname="countyreal"`  

* There needs to be a time variable.  In `mpdta`, it is the variable `year`.  The name of this variable is passed to methods in the `did` package by setting, for example, `tname="year"`

* In this application, the outcome is `lemp`.  The name of this variable is passed to methods in the `did` package by setting, for example, `yname="lemp"`

* There needs to be a group variable.  In `mpdta`, it is the variable `first.treat`.  This is the time period when an individual first becomes treated.   *For individuals that are never treated, this variable should be set equal to 0.*  The name of this variable is passed to methods in the `did` package by setting, for example, `first.treat.name="first.treat"`

* The `did` package allows for incorporating covariates so that the parallel trends assumption holds only after conditioning on these covariates.  In `mpdta`, `lpop` is the log of county population.  The `did` package requires that covariates be time-invariant.  For time varying covariates like county population, the `did` package sets the value of the covariate to be equal to the value of the covariate in the first time period.  Covariates are passed as a formula to the `did` package by setting, for example, `xformla=~lpop`.  For estimators under unconditional parallel trends, the `xformla` argument can be left blank or can be set as `xformla=~1` to only include a constant.

### The Effect of the Minimum Wage on Youth Employment

Next, we walk through a straightforward, but realistic way to use the `did` package to carry out an application.

**Side Comment:** This is just an example of how to use our method in a semi-realistic setup.  To really evaluate the effect of the minimum wage on teen employment, one would need to be more careful along a number of dimensions.


We'll consider two cases.  For the first case, we won't condition on any covariates.  For the second, we'll condition on the log of county population (in a "real" application, one might want to condition on more covariates).

```{r}
# estimate group-time average treatment effects without covariates
# set bstrap=TRUE and cband=TRUE to get uniform confidence bands
mw.attgt <- att_gt(yname="lemp",
                   first.treat.name="first.treat",
                   idname="countyreal",
                   tname="year",
                   xformla=~1,
                   data=mpdta,
                   bstrap=TRUE,
                   cband=TRUE,
                   printdetails=FALSE
                   )

# summarize the results
summary(mw.attgt)

# plot the results
# set ylim so that all plots have the same scale along y-axis
ggdid(mw.attgt, ylim=c(-.3,.3))
```

There are a few things to notice in this case

* There does not appear to be much evidence against the parallel trends assumption.  One fails to reject using the Wald test reported in `summary`; likewise the uniform confidence bands cover 0 in all pre-treatment periods.

* There is some evidence of negative effects of the minimum wage on employment.  Two group-time average treatment effects are negative and statistically different from 0.  These results also suggest that it may be helpful to aggregate the group-time average treatment effects.

```{r}
# aggregate the group-time average treatment effects
mw.dyn <- aggte(mw.attgt, type="dynamic")
summary(mw.dyn)
ggdid(mw.dyn, ylim=c(-.3,.3))
```

These continue to be uniform confidence bands for dynamic effects.  The results are broadly similar to the ones from the group-time average treatment effects:  one fails to reject parallel trends in pre-treatment periods and it looks like somewhat negative effects of the minimum wage on youth employment.

One potential issue with these dynamic effect estimators is that the composition of the groups changes with different lengths of exposure in the event study plots.  For example, for the group of states who increased their minimum wage in 2007, we only see the effect of the minimum wage on impact, but we do not observe any dynamic effects.  If the effects of the minimum wage are systematically different across groups (here, there is not much evidence of this as the effect for all groups seems to be close to 0 on impact and perhaps becoming more negative over time), then this can lead to confounding dynamics and selective treatment timing among different groups.  One way to combat this is to balance the sample by (i) only including groups that are exposed to the treatment for at least a certain number of time periods and (ii) only look at dynamic effects in those time periods.  In the `did` package, one can do this by specifying the `balance.e` option.  Here, we set `balance.e=1` -- what this does is to only consider groups of states that are treated in 2004 and 2006 (and drops the group treated in 2007) and then only looks at dynamics on impact and in the next period following raising the minimum wage.

```{r}
mw.dyn.balance <- aggte(mw.attgt, type="dynamic", balance.e=1)
summary(mw.dyn.balance)
ggdid(mw.dyn.balance, ylim=c(-.3,.3))
```

Finally, we can run all the same results including a covariate.  In this application the results turn out to be nearly identical, and here we provide just the code for estimating the group-time average treatment effects while including covariates.  The other steps are otherwise the same.

```{r eval=FALSE}
mw.attgt.X <- att_gt(yname="lemp",
                   first.treat.name="first.treat",
                   idname="countyreal",
                   tname="year",
                   xformla=~lpop,
                   data=mpdta,
                   bstrap=TRUE,
                   cband=TRUE,
                   printdetails=FALSE
                   )
```


## Common Issues with the did package

We update this section with common issues that people run into when using the `did` package.  Please feel free to contact us with questions or comments.

* In the case with panel data, the `did` package "balances" the panel by dropping individual units which do not have observations in each time period.

* The `did` package is only built to handle staggered treatment adoption designs.  This means that once an individual becomes treated, they remain treated in all subsequent periods.

